### 2024-09-25 paper/paper.qmd
Finished graphs and updated paper

---
title: "Toronto Homicides and Shootings (To be Changed)"
subtitle: "An analysis of police presence across Toronto neighbourhoods in relation to homicide and shooting cases (2018-2023)"
author: Emily Su
thanks: "Code and data are available at: https://github.com/moonsdust/toronto-homicides-shootings."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format:
  pdf:
    toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library("tidyverse") # Contains data-related packages
library("knitr") # To make tables
library("opendatatoronto")
library("ggplot2") # To make graphs
library("sf") # To make maps
library("geojsonsf") # To read geojson files
```

```{r}
#| include: false
#| warning: false
#| message: false

# Read in data
# Police data 
cleaned_police_location <- geojson_sf('../data/analysis_data/cleaned_police_location.geojson')
# Crime data
cleaned_crime_data <- geojson_sf('../data/analysis_data/cleaned_crime_data.geojson')
# Ward data 
cleaned_ward_data <- read_csv('../data/analysis_data/cleaned_ward_data.csv', show_col_types = FALSE)
```

# Introduction

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}
## Methodology

## Data Source and Measurements

## Variables of Interest

# Results {#sec-results}
## Number of Homicide and Shooting Cases in Toronto from 2018 to 2023 in Toronto
```{r}
#| include: false
#| warning: false
#| message: false
# Add up total number of cases per neighbourhood
crime_per_neighbourhood <- 
  cleaned_crime_data |>
  # Calculate the total sum of cases across all neighbourhoods 
  mutate(
    num_cases_total = sum(num_of_cases)
  ) |>
  select(neighbourhood, homicide_or_shooting, num_of_cases, num_cases_total) |>
  # Groups data by type of crime, and neighbourhood
  group_by(neighbourhood, homicide_or_shooting) |>
  # Create new columns for number of cases for homicides and shootings per neighbourhood. Also calculate the proportion as well. 
  mutate (
    num_cases_total_per_neighbourhood_per_crime = sum(num_of_cases),
    num_cases_total_per_neighbourhood_per_crime_proportion = round(num_cases_total_per_neighbourhood_per_crime / num_cases_total, 3)
  ) |>
  ungroup() |>
  group_by(neighbourhood) |> 
  mutate(
    num_cases_total_per_neighbourhood = sum(num_of_cases),
    num_cases_total_per_neighbourhood_proportion = round(num_cases_total_per_neighbourhood / num_cases_total, 3)
  ) |>
  # Obtains unique rows
  unique() |>
  # Arrange based on the total number of cases per neighbourhood
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  tibble() |>
  select(neighbourhood, homicide_or_shooting, num_of_cases, num_cases_total, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_proportion, num_cases_total_per_neighbourhood_per_crime_proportion)

# Add up total number of cases each year across all neighbourhoods
yearly_crime <- 
  cleaned_crime_data |>
  # Calculate the total sum of cases across all neighbourhoods 
  mutate(
    num_cases_total = sum(num_of_cases)
  ) |>
  # Group data by year, type of crime, number of total cases, and number of cases
  select(year, homicide_or_shooting, num_cases_total, num_of_cases) |>
  group_by(year, homicide_or_shooting) |>
  unique() |>
  # Create new columns for number of cases per year for homicides and shootings across ALL neighbourhood. Also calculate the proportion as well. 
  mutate (
    yearly_num_cases_total = sum(num_of_cases),
    yearly_num_cases_total_proportion = round(yearly_num_cases_total / num_cases_total, 3)
  ) |>
  tibble() |>
  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_proportion) |>
  unique() |>
  # Arrange based on the total number of cases per year 
  arrange(desc(yearly_num_cases_total)) |>
  ungroup()
```

```{r}
#| label: tbl-totalcasestable
#| fig-cap:  Number of Homicide and Shooting Cases in Toronto from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
yearly_crime |>
  select(year, homicide_or_shooting, yearly_num_cases_total, yearly_num_cases_total_proportion) |>
  # Arrange based on year
  arrange(year) |>
  unique() |>
  rename(
    "Type of crime" = homicide_or_shooting,
    "Total number of cases yearly" = yearly_num_cases_total,
    "Proportion of cases yearly" = yearly_num_cases_total_proportion,
    "Year" = year
  ) |>
  kable()
```

```{r}
#| label: fig-totalcaseshistogram
#| fig-cap:  Number of Homicide and Shooting Cases in Toronto from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

yearly_crime |> 
  ggplot(mapping = aes(x = year, y = yearly_num_cases_total, fill = homicide_or_shooting)) +
  geom_bar(stat="identity", position = "dodge2") +
  labs(x = "Year", y = "Number of Cases", fill = "Type of Crime") +
  theme_minimal() + 
  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7)) + scale_fill_viridis_d()
```

```{r}
#| label: tbl-totalcasesstats
#| tbl-cap: Minimum, quartiles, median, and maximum of the number of homicides and shootings in Toronto from 2018 to 2023
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

# Creates table
yearly_crime |>
  select(yearly_num_cases_total) |>
  rename(
    "Total number of cases" = yearly_num_cases_total
  ) |>
  summary() |>
  kable()
```

```{r}
#| label: tbl-totalshootingsstats
#| tbl-cap: Minimum, quartiles, median, and maximum of the number of shootings in Toronto from 2018 to 2023
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

# Creates table
yearly_crime |>
  filter(homicide_or_shooting == "shooting") |>
  select(yearly_num_cases_total) |>
  rename(
    "Total number of cases" = yearly_num_cases_total
  ) |>
  summary() |>
  kable()
```

```{r}
#| label: tbl-totalhomicidesstats
#| tbl-cap: Minimum, quartiles, median, and maximum of the number of homicides in Toronto from 2018 to 2023
#| echo: false # Code will run and its output will show but the code itself will not be included in the doc
#| warning: false # Warnings are not included 
#| message: false # Messages are not included

# Creates table
yearly_crime |>
  filter(homicide_or_shooting == "homicide") |>
  select(yearly_num_cases_total) |>
  rename(
    "Total number of cases" = yearly_num_cases_total
  ) |>
  summary() |>
  kable()
```

## Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings

```{r}
#| label: tbl-highestcasesshootingstable
#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Shootings from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "shooting") |>
  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
  select(neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

```{r}
#| label: tbl-highestcaseshomicidestable
#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "homicide") |>
  arrange(desc(num_cases_total_per_neighbourhood_per_crime)) |>
  select(neighbourhood, num_cases_total_per_neighbourhood_per_crime, num_cases_total_per_neighbourhood_per_crime_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_per_crime_proportion,
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

```{r}
#| label: tbl-highestcasestable
#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings Combined from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

# Creates table
crime_per_neighbourhood |>
  filter(homicide_or_shooting == "homicide") |>
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  select(neighbourhood, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_proportion) |>
  unique() |>
  rename(
    "Total number of cases (2018 to 2023)" = num_cases_total_per_neighbourhood,
    "Proportion of cases (2018 to 2023)" = num_cases_total_per_neighbourhood_proportion,
    "Neighbourhood" = neighbourhood
  ) |>
  head(6) |>
  kable()
```

```{r}
#| label: fig-highestcaseshistogram
#| fig-cap:  Six Toronto Neighbourhoods with the Highest Cases of Homicides and Shootings from 2018 to 2023
#| echo: false
#| warning: false
#| message: false

crime_per_neighbourhood |> 
  select(neighbourhood, homicide_or_shooting, num_cases_total_per_neighbourhood, num_cases_total_per_neighbourhood_per_crime) |>
  arrange(desc(num_cases_total_per_neighbourhood)) |>
  select(-c(num_cases_total_per_neighbourhood)) |>
  unique() |>
  head(12) |>
  ggplot(mapping = aes(x = neighbourhood, y = num_cases_total_per_neighbourhood_per_crime, fill = homicide_or_shooting)) +
  geom_bar(stat="identity", position = "dodge2") +
  labs(x = "Year", y = "Number of Cases", fill = "Type of Crime") +
  theme_minimal() + 
  # Reference https://stackoverflow.com/questions/1330989/rotating-and-spacing-axis-labels-in-ggplot2 to figure out how to rotate text along x axis.
  theme(legend.position = "bottom", plot.title = element_text(size = 9, face = "bold"), legend.title = element_text(size = 9), legend.text = element_text(size = 7), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))  + scale_fill_viridis_d()
```

## Police Facility Locations in relation to the Number of Homicides and Shootings from 2018 to 2023 in Toronto

```{r}
#| label: fig-homicides-map
#| fig-cap: Location of police facilities in relation to Toronto neighbourhoods with the highest number of homicides from 2018 to 2023.
#| echo: false
#| warning: false
#| message: false
# Reference the following for code:
# https://cengel.github.io/R-spatial/mapping.html
ggplot() + 
    # To help convert between GeoJSON to sf  https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html
  # Fill in area regarding the number of cases 
  geom_sf(data = cleaned_crime_data |> filter(homicide_or_shooting == "homicide"), aes(fill = num_of_cases, geometry = geometry)) +
  # Changes the font of the title and removes the x and y axis 
  # Reference to figured out how to add labels to maps https://tmieno2.github.io/R-as-GIS-for-Economists/create-maps.html
  theme_minimal() + 
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "Number of homicides across Toronto neighbourhoods (2018 to 2023)",
    fill = "Number of cases"
  ) +
  geom_sf_text(
    data = cleaned_crime_data |> filter(num_of_cases >= 14),
    aes(label = hood_id),
    check_overlap = TRUE,
    size = 2, 
    color = "white"
    ) +
    theme(plot.title = element_text(size = 9, face = 'bold'), 
          plot.subtitle = element_text(size = 9),
          rect = element_blank(), 
          axis.text.x = element_blank(), 
          axis.text.y = element_blank(), 
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.ticks = element_blank(), 
          legend.title = element_text(size = 8),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
          ) +
  # Add in police locations
  geom_point(data = cleaned_police_location,
    aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
  ) + labs(color = "Location")

```

```{r}
#| label: fig-shootings-map
#| fig-cap: Location of police facilities in relation to Toronto neighbourhoods with the highest number of shootings from 2018 to 2023.
#| echo: false
#| warning: false
#| message: false
# Reference the following for code:
# https://cengel.github.io/R-spatial/mapping.html
ggplot() + 
    # To help convert between GeoJSON to sf  https://cran.r-project.org/web/packages/geojsonsf/vignettes/geojson-sf-conversions.html
  # Fill in area regarding the number of cases 
  geom_sf(data = cleaned_crime_data |> filter(homicide_or_shooting == "shooting"), aes(fill = num_of_cases, geometry = geometry)) +
  # Changes the font of the title and removes the x and y axis 
  # Reference to figured out how to add labels to maps https://tmieno2.github.io/R-as-GIS-for-Economists/create-maps.html
  theme_minimal() + 
  labs(
    x = "Longitude",
    y = "Latitude",
    title = "Number of shootings across Toronto neighbourhoods (2018 to 2023)",
    fill = "Number of cases"
  ) +
  geom_sf_text(
    data = cleaned_crime_data |> filter(num_of_cases >= 14),
    aes(label = hood_id),
    check_overlap = TRUE,
    size = 2, 
    color = "white"
    ) +
    theme(plot.title = element_text(size = 9, face = 'bold'), 
          plot.subtitle = element_text(size = 9),
          rect = element_blank(), 
          axis.text.x = element_blank(), 
          axis.text.y = element_blank(), 
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.ticks = element_blank(), 
          legend.title = element_text(size = 8),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()
          ) +
  # Add in police locations
  geom_point(data = cleaned_police_location,
    aes(geometry = geometry, color = "Police Facility"), stat = "sf_coordinates"
  ) + labs(color = "Location")

```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Areas of improvement

Weaknesses and next steps should also be included.

## Next steps

\newpage

\appendix

# Appendix {-}

## Acknowledgments 
We would like the acknowledge @citetellingstorieswithdata for some R code that 
was used to produce the tables and graphs.

## Note on Reproducing 
To reproduce the results in the paper, first run the scripts found in the scripts folder of the GitHub repository corresponding to the paper starting with the script, 00-install_packages.R to install the necessary packages.

## Code styling 
Code written in the scripts was checked and styled with lintr [@citelintr] and styler [@citestyler].

## Additional Figures

## Additional Tables

\newpage

# References